Foregrounds in Wide--Field Redshifted 21~cm Power Spectra

### 2015-02-23 ms.tex
Remove all bold face fonts
from 37bab06:
============

In ms.tex,

(*) Remove all bold face fonts

Detection of 21~cm emission of H{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission. {\bf We investigate the signatures of wide--field measurements and an all--sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs. We demonstrate interferometric measurements are inherently sensitive to all scales, including the largest angular scales, owing to the nature of wide--field measurements. These wide--field effects are generic to all observations but antenna shapes impact their amplitudes substantially. A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array}. Comparing data from recent Murchison Widefield Array observations, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. We identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf that usually represent structures on small scales}. For signals entering through the primary field of view, compact emission dominates the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier delay space. Based on these results, we propose that selective down--weighting of data based on antenna spacing and time can mitigate foreground contamination substantially by a factor $\sim 100$ with negligible loss of sensitivity.

@@

Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}. Numerous {\bf pathfinders and} precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}. {\bf The Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) is currently under construction using new insights gained with the MWA and PAPER}.

A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12b,dil13,pob13}. {\bf Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity}.

Considerable effort is being made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}. {\bf A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'}. The conservative foreground strategy, referred to as {\it avoidance}, that has developed alongside this work is to discard $k$--modes which could be contaminated \citep[e.g.,][]{par14}. The more aggressive alternative is to subtract a sky model and regain access to modes that would be discarded by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and proposing a scheme for down--weighting these components.

Foregrounds with intrinsic deviations from spectral smoothness, instruments with high chromaticity, polarization leakage, calibration errors, or approximations in power spectrum analyses can contaminate the true EoR H{\sc i} power spectrum. Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity. To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using a {\it \textbf{per-baseline}} approach of \citet{par12b}. This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.

@@

{\bf In \S\ref{sec:delay-spectrum} we provide an overview of the delay spectrum technique. We investigate signatures generic to all wide--field measurements of EoR power spectra in \S\ref{sec:wide-field}. In \S\ref{sec:sim}, we present the foreground model and a variety of instrument models to rank antenna shapes based on foreground contamination. In \S\ref{sec:MWA}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are compared with the data and analyzed for foreground signatures. We report two important findings: foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique that minimizes foreground contamination using prior knowledge of the sky to down--weight adversely contaminated baselines}. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

Interferometer array data known as {\it visibilities}, $V_b(f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis. If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_\textrm{i}(f)$ denotes instrumental bandpass weights, then $V_b(f)$ can be written as:
\begin{align}\label{eqn:obsvis}
  V_b(f) &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
\end{align}
where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector. This equation is valid {\bf in general, including wide--field measurements, and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}}.

{\bf The {\it delay spectrum}, $\tilde{V}_b(\tau)$, is defined as} the inverse Fourier transform of $V_b(f)$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  \tilde{V}_b(\tau) &\equiv \int V_b(f)\,W(f)\,e^{i2\pi f\tau}\,\dif f,
\end{align}
where, $W(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
\begin{equation}\label{eqn:delay}
  \tau = \frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}.
\end{equation}
The delay transform conventions used in this paper are described in appendix~\ref{sec:delay-conventions}. $\tilde{V}_b(\tau)$ is expressed in observer's units as Jy~Hz.

@@

{\bf We give a brief overview of some parameters of Fourier space which is generic to all experiments that use a similar approach}. Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of redshifted H{\sc i} observations are calculated. $|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction {\bf (tangent plane to the celestial sphere)}, form the $x$--axis. $\tau$ and $k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them. These boundaries, called horizon delay limits \citep{ved12,par12b}, are shown by solid lines.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
% \includegraphics[width=\linewidth]{f1.eps}
% \includegraphics[width=\linewidth]{fig2.eps}
% \includegraphics[width=\linewidth]{figure2.eps}
\includegraphics[width=\linewidth]{figure1.eps}
\caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are calculated. The $x$--axis is denoted by $|\boldsymbol{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber). The $y$--axis denoted by $\tau$ (delay) or $k_\parallel$ (line--of--sight wavenumber). Here, $k_\perp$ and $k_\parallel$ are obtained for a frequency of 185~MHz. The dark shaded region is referred to as the {\it foreground wedge} where smooth spectrum foregrounds reside. Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits. Narrow extensions of the {\it wedge} (white unshaded strips) are caused by convolution with the {\bf instrument's spectral transfer function}. Regions excluding the {\it wedge} are expected to be relatively free of foreground contamination {\bf and are generally referred to as the {\it EoR window}. There are undesirable grating responses (dot--dashed lines) specific to the MWA. Hence, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the MWA {\it EoR window}}. \label{fig:fourier-space}}
\end{figure}

The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis \citep{par12b,thy13}. The width of this narrow strip is inversely proportional to the operating bandwidth. The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades. {\bf In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$. Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays. Thus the regions of interest for EoR studies relying on {\it avoidance} strategy are just beyond the horizon delay limits (dashed lines) on short baselines, marked as regions of {\it maximal EoR sensitivity}}.

{\bf In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s. Thus the MWA {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade}.

@@

{\bf In our present study, we use a latitude of -26.701\arcdeg~ and an antenna layout identical to that of the MWA \citep{bea12} for the observatory. The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km. Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization). Their deviation from coplanarity is negligible. For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$. Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis. We choose an observing frequency of 185~MHz ($z\simeq 6.68$) and a flat passband of width $\Delta B = 30.72$~MHz to roughly match those of ongoing MWA EoR observations (discussed in detail in \S\ref{sec:MWA})}.

@@

The model described above is primarily a model of the diffuse foreground sky. While it contains faint compact emission blended in with the diffuse emission, {\bf bright point sources have been removed \citep{deo08}}. In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and include point sources brighter than this threshold. Slightly different criteria are in common use in radio astronomy to estimate radio source confusion \citep[see Appendix of][and references therein]{thy13}. For an angular resolution of 0.85\arcdeg, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy. Other liberal criteria that yield a lower threshold carry a greater risk of double--counting point sources which might be already blended in with the diffuse sky model.

@@

\begin{figure*}[htb]
\centering
% \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure5a.eps}} \\
% \subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure5b.eps}}
\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure4a.eps}} \\
\subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure4b.eps}}
\caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith. Antenna shapes used are: dipole (left), phased array (middle), and dish (right). Refer to \S\ref{sec:instrument_model} for details of the antenna models. The strength and occupancy of the power patterns are correlated with those of delay power spectra. White dotted lines in the delay power spectra mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The {\it foreground wedge} and the {\it EoR window} are most severely contaminated in the case of the dipole while it is the least for the dish. The phased array has intermediate levels of contamination. Foreground power close to the horizon delay limits in all three cases are significant even on long baselines. The foreshortening of baselines towards the horizon makes them sensitive to foreground emission on large size scales. The amplitude of this feature strongly depends on the shape of the antenna element. It is highest for a dipole (which has a strong response near the horizon) and least for a dish.}
\label{fig:aperture-shapes}
\end{figure*}

@@

The MWA passband of width $\Delta B=30.72$~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter. The coarse channel shape is obtained using an 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$. {\bf Each of these coarse bands consists of 32 fine channels of width 40~kHz each}. After correcting for the shape of these coarse channels, the fine channels at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level. % Any spectral flags, or other known weights constitutes $W_\textrm{i}(f)$ in equation~\ref{eqn:obsvis}.

@@

{\bf The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{bow06,bea13,thy13}. Over 600~hours have been currently observed using the MWA, targeting science objectives outlined in \citet{bow13}.}

@@

The observations {\bf used} here were acquired on 2013~August~23. We have chosen two sections of duration 112 seconds each from this night for detailed study. These were chosen to provide a selection of possible foreground and instrumental conditions. As an example of a nominal observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hours from zenith. This pointing has a significantly higher secondary lobe structure and is observed when the bright galactic center is well above the horizon.

@@

{\bf We estimate thermal noise, $\Delta V$, in the observed visibilities, $V_b(f)$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained from data after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
\begin{align}\label{eqn:Tsys}
  \Delta \tilde{V} &= \sqrt{N_\textrm{ch}}\,\Delta V \Delta f, \quad\textrm{and} \\
  \Delta V &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}}{A_\textrm{e}\sqrt{2\,\Delta f\,\Delta t}},
\end{align}
where, $\Delta f = 80$~kHz, $\Delta t = 112$~s, and $N_\textrm{ch}=\Delta B/\Delta f$ is the number of frequency channels. The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus yields a robust estimate of $T_\textrm{sys}$. We find the average system temperature to be $\sim 95$~K. Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.}

@@

With the aforementioned foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}. Figure~\ref{fig:fhd-sim-comparison} shows the {\bf delay power spectra} from {\it off--zenith} and {\it zenith} pointings obtained from MWA observations and modeling. Notice the qualitative agreement of amplitude and structure between the two. The Galactic center and the Galactic plane visible in the {\it off--zenith} pointing make it appear brighter in the {\it foreground wedge} as a branch with $\tau<0$.

@@

We also simulated delay spectra after assigning spectral indices drawn randomly from a {\it gaussian} distribution with a mean of $\langle\alpha_\textrm{sp}\rangle=-0.83$ and a HWHM of $\Delta\alpha_\textrm{sp}=0.35$ to the point sources in our compact foreground model. These simulations typically yielded a median absolute deviation of $\approx 0.29$ for $\log_{10}\rho$ indicating fractional differences of $\sim$~95\% {\bf between different realizations}. This demonstrates that a fractional deviation of $\sim$~90\% observed between data and simulations is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account.

>>>>

Detection of 21~cm emission of H{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission. We investigate the signatures of wide--field measurements and an all--sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs. We demonstrate interferometric measurements are inherently sensitive to all scales, including the largest angular scales, owing to the nature of wide--field measurements. These wide--field effects are generic to all observations but antenna shapes impact their amplitudes substantially. A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array. Comparing data from recent Murchison Widefield Array observations, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. We identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings that usually represent structures on small scales. For signals entering through the primary field of view, compact emission dominates the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier delay space. Based on these results, we propose that selective down--weighting of data based on antenna spacing and time can mitigate foreground contamination substantially by a factor $\sim 100$ with negligible loss of sensitivity.

@@

Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}. Numerous pathfinders and precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}. The Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) is currently under construction using new insights gained with the MWA and PAPER.

A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12b,dil13,pob13}. Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity.

Considerable effort is being made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}. A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'. The conservative foreground strategy, referred to as {\it avoidance}, that has developed alongside this work is to discard $k$--modes which could be contaminated \citep[e.g.,][]{par14}. The more aggressive alternative is to subtract a sky model and regain access to modes that would be discarded by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and proposing a scheme for down--weighting these components.

Foregrounds with intrinsic deviations from spectral smoothness, instruments with high chromaticity, polarization leakage, calibration errors, or approximations in power spectrum analyses can contaminate the true EoR H{\sc i} power spectrum. Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity. To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using a {\it per-baseline} approach of \citet{par12b}. This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.

@@

In \S\ref{sec:delay-spectrum} we provide an overview of the delay spectrum technique. We investigate signatures generic to all wide--field measurements of EoR power spectra in \S\ref{sec:wide-field}. In \S\ref{sec:sim}, we present the foreground model and a variety of instrument models to rank antenna shapes based on foreground contamination. In \S\ref{sec:MWA}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are compared with the data and analyzed for foreground signatures. We report two important findings: foregrounds that most severely obscure the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique that minimizes foreground contamination using prior knowledge of the sky to down--weight adversely contaminated baselines. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

Interferometer array data known as {\it visibilities}, $V_b(f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis. If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_\textrm{i}(f)$ denotes instrumental bandpass weights, then $V_b(f)$ can be written as:
\begin{align}\label{eqn:obsvis}
  V_b(f) &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
\end{align}
where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector. This equation is valid in general, including wide--field measurements, and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}.

The {\it delay spectrum}, $\tilde{V}_b(\tau)$, is defined as the inverse Fourier transform of $V_b(f)$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  \tilde{V}_b(\tau) &\equiv \int V_b(f)\,W(f)\,e^{i2\pi f\tau}\,\dif f,
\end{align}
where, $W(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
\begin{equation}\label{eqn:delay}
  \tau = \frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}.
\end{equation}
The delay transform conventions used in this paper are described in appendix~\ref{sec:delay-conventions}. $\tilde{V}_b(\tau)$ is expressed in observer's units as Jy~Hz.

@@

We give a brief overview of some parameters of Fourier space which is generic to all experiments that use a similar approach. Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of redshifted H{\sc i} observations are calculated. $|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction (tangent plane to the celestial sphere), form the $x$--axis. $\tau$ and $k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them. These boundaries, called horizon delay limits \citep{ved12,par12b}, are shown by solid lines.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
% \includegraphics[width=\linewidth]{f1.eps}
% \includegraphics[width=\linewidth]{fig2.eps}
% \includegraphics[width=\linewidth]{figure2.eps}
\includegraphics[width=\linewidth]{figure1.eps}
\caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are calculated. The $x$--axis is denoted by $|\boldsymbol{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber). The $y$--axis denoted by $\tau$ (delay) or $k_\parallel$ (line--of--sight wavenumber). Here, $k_\perp$ and $k_\parallel$ are obtained for a frequency of 185~MHz. The dark shaded region is referred to as the {\it foreground wedge} where smooth spectrum foregrounds reside. Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits. Narrow extensions of the {\it wedge} (white unshaded strips) are caused by convolution with the instrument's spectral transfer function. Regions excluding the {\it wedge} are expected to be relatively free of foreground contamination and are generally referred to as the {\it EoR window}. There are undesirable grating responses (dot--dashed lines) specific to the MWA. Hence, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the MWA {\it EoR window}. \label{fig:fourier-space}}
\end{figure}

The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis \citep{par12b,thy13}. The width of this narrow strip is inversely proportional to the operating bandwidth. The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades. In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$. Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays. Thus the regions of interest for EoR studies relying on {\it avoidance} strategy are just beyond the horizon delay limits (dashed lines) on short baselines, marked as regions of {\it maximal EoR sensitivity}.

In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s. Thus the MWA {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade.

@@

In our present study, we use a latitude of -26.701\arcdeg~ and an antenna layout identical to that of the MWA \citep{bea12} for the observatory. The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km. Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization). Their deviation from coplanarity is negligible. For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$. Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis. We choose an observing frequency of 185~MHz ($z\simeq 6.68$) and a flat passband of width $\Delta B = 30.72$~MHz to roughly match those of ongoing MWA EoR observations (discussed in detail in \S\ref{sec:MWA}).

@@

The model described above is primarily a model of the diffuse foreground sky. While it contains faint compact emission blended in with the diffuse emission, bright point sources have been removed \citep{deo08}. In order to supplement it with missing bright compact emission, we use classical radio source confusion estimates to determine the nominal flux density threshold and include point sources brighter than this threshold. Slightly different criteria are in common use in radio astronomy to estimate radio source confusion \citep[see Appendix of][and references therein]{thy13}. For an angular resolution of 0.85\arcdeg, using a conservative `$S_\textrm{c}=5\sigma_\textrm{c}$' criterion, we determine the flux density threshold to be $\approx 10$~Jy. Other liberal criteria that yield a lower threshold carry a greater risk of double--counting point sources which might be already blended in with the diffuse sky model.

@@

\begin{figure*}[htb]
\centering
% \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure5a.eps}} \\
% \subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure5b.eps}}
\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure4a.eps}} \\
\subfloat[][Delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure4b.eps}}
\caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith. Antenna shapes used are: dipole (left), phased array (middle), and dish (right). Refer to \S\ref{sec:instrument_model} for details of the antenna models. The strength and occupancy of the power patterns are correlated with those of delay power spectra. White dotted lines in the delay power spectra mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The {\it foreground wedge} and the {\it EoR window} are most severely contaminated in the case of the dipole while it is the least for the dish. The phased array has intermediate levels of contamination. Foreground power close to the horizon delay limits in all three cases are significant even on long baselines. The foreshortening of baselines towards the horizon makes them sensitive to foreground emission on large size scales. The amplitude of this feature strongly depends on the shape of the antenna element. It is highest for a dipole (which has a strong response near the horizon) and least for a dish.}
\label{fig:aperture-shapes}
\end{figure*}

@@

The MWA passband of width $\Delta B=30.72$~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter. The coarse channel shape is obtained using an 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$. Each of these coarse bands consists of 32 fine channels of width 40~kHz each. After correcting for the shape of these coarse channels, the fine channels at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level. % Any spectral flags, or other known weights constitutes $W_\textrm{i}(f)$ in equation~\ref{eqn:obsvis}.

@@

The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{bow06,bea13,thy13}. Over 600~hours have been currently observed using the MWA, targeting science objectives outlined in \citet{bow13}.

@@

The observations used here were acquired on 2013~August~23. We have chosen two sections of duration 112 seconds each from this night for detailed study. These were chosen to provide a selection of possible foreground and instrumental conditions. As an example of a nominal observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hours from zenith. This pointing has a significantly higher secondary lobe structure and is observed when the bright galactic center is well above the horizon.

@@

We estimate thermal noise, $\Delta V$, in the observed visibilities, $V_b(f)$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained from data after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
\begin{align}\label{eqn:Tsys}
  \Delta \tilde{V} &= \sqrt{N_\textrm{ch}}\,\Delta V \Delta f, \quad\textrm{and} \\
  \Delta V &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}}{A_\textrm{e}\sqrt{2\,\Delta f\,\Delta t}},
\end{align}
where, $\Delta f = 80$~kHz, $\Delta t = 112$~s, and $N_\textrm{ch}=\Delta B/\Delta f$ is the number of frequency channels. The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus yields a robust estimate of $T_\textrm{sys}$. We find the average system temperature to be $\sim 95$~K. Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.

@@

With the aforementioned foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}. Figure~\ref{fig:fhd-sim-comparison} shows the delay power spectra from {\it off--zenith} and {\it zenith} pointings obtained from MWA observations and modeling. Notice the qualitative agreement of amplitude and structure between the two. The Galactic center and the Galactic plane visible in the {\it off--zenith} pointing make it appear brighter in the {\it foreground wedge} as a branch with $\tau<0$.

@@

We also simulated delay spectra after assigning spectral indices drawn randomly from a {\it gaussian} distribution with a mean of $\langle\alpha_\textrm{sp}\rangle=-0.83$ and a HWHM of $\Delta\alpha_\textrm{sp}=0.35$ to the point sources in our compact foreground model. These simulations typically yielded a median absolute deviation of $\approx 0.29$ for $\log_{10}\rho$ indicating fractional differences of $\sim$~95\% between different realizations. This demonstrates that a fractional deviation of $\sim$~90\% observed between data and simulations is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account.
